Estimating SARS-CoV-2 transmission in England from randomly sampled prevalence surveys

### 2024-03-25 ATACCC-appendix-original-analysis.tex
Tweak final plots / captions

\begin{figure}
    \centering
    \vspace*{-3cm}
    \makebox[\textwidth][c]{\includegraphics[scale=0.9]{ATACCC/appendix/goodness_of_fit}}
    \caption[Goodness-of-fit]{Predictive median (solid line), 50\% and 95\% predictive intervals (light and dark shading respectively) for test results (ignoring false negatives). Points show actual observations.}
    \label{fig:paper:goodness_of_fit}
\end{figure}

@@

This study presents estimates of viral load dynamics based on a general population sample with daily swabbing.
Our results show a peak Ct value around 5 Ct lower than \textcite{kisslerViral}, that is a higher viral load, although occurring at a similar time (see \cref{tab:paper:compare}) and a longer time from first detection to peak viral load than \textcite{jonesEstimating}.
This could be because of the different populations in each study. \Textcite{kisslerViral}'s data is collected from longitudinal swabbing of individuals professionally involved in the National Basketball Association (NBA).
Of the individuals who were included in their analysis, around half were in professional athletes, likely in peak physical condition, and 90\% were male.
Our data is collected from the general population who were reported as contacts of individuals detected through surveillance open to all members of the public.
Therefore, our sample is likely to be, on average, in worse physical condition which may lead to a higher viral load (lower Ct).
Another possible difference could be the lab the tests were performed in.
Ct values are not fully comparable between different laboratories due to different testing methodologies~\autocites{dahdouhCt,hanRTPCR}.

>>>>

\begin{figure}
    \thisfloatpagestyle{empty}
    \vspace*{-1cm}
    \makebox[\textwidth][c]{\includegraphics[scale=0.9]{ATACCC/appendix/goodness_of_fit}}
    \caption[Goodness-of-fit]{Predictive median (solid line), 50\% and 95\% predictive intervals (light and dark shading respectively) for test results (ignoring false negatives). Points show actual observations.}
    \label{fig:paper:goodness_of_fit}
\end{figure}

@@

This study presents estimates of viral load dynamics based on a general population sample with daily swabbing.
Our results show a peak Ct value around 5 Ct lower than \textcite{kisslerViral}, that is a higher viral load, although occurring at a similar time (see \cref{tab:paper:compare}) and a longer time from first detection to peak viral load than \textcite{jonesEstimating}.
This could be because of the different populations in each study. \Textcite{kisslerViral}'s data are collected from longitudinal swabbing of individuals professionally involved in the National Basketball Association (NBA).
Of the individuals who were included in their analysis, around half were in professional athletes, likely in peak physical condition, and 90\% were male.
Our data are collected from the general population who were reported as contacts of individuals detected through surveillance open to all members of the public.
Therefore, our sample is likely to be, on average, in worse physical condition which may lead to a higher viral load (lower Ct).
Another possible difference could be the lab the tests were performed in.
Ct values are not fully comparable between different laboratories due to different testing methodologies~\autocites{dahdouhCt,hanRTPCR}.

### 2024-03-25 biology-data.tex

The household-based design of the CIS means that the data is clustered.
In particular, individuals in the same household are likely to transmit to each other and hence have correlated test results.
Other behavioural or genetic factors could also contribute to individuals in the same household being more similar than randomly selected individuals in the population.
The positivity analyses use a beta-binomial likelihood (defined in \cref{E-distributions}) to account for this.

@@

For privacy reasons, the CIS data is stored in the ONS's SRS (Secure Research Service)~\autocite{onsSRS}.
The SRS is a Trusted Research Environment (TRE) that allows researchers to access sensitive data in a secure environment.
The SRS gives accredited researchers working on projects in the public interest access to granular data which cannot be legally shared publicly.

>>>>

The household-based design of the CIS means that the data are clustered.
In particular, individuals in the same household are likely to transmit to each other and hence have correlated test results.
Other behavioural or genetic factors could also contribute to individuals in the same household being more similar than randomly selected individuals in the population.
The positivity analyses use a beta-binomial likelihood (defined in \cref{E-distributions}) to account for this.

@@

For privacy reasons, the CIS data are stored in the ONS's SRS (Secure Research Service)~\autocite{onsSRS}.
The SRS is a Trusted Research Environment (TRE) that allows researchers to access sensitive data in a secure environment.
The SRS gives accredited researchers working on projects in the public interest access to granular data which cannot be legally shared publicly.

### 2024-03-25 cis-perfect-testing.tex

The CIS tests each individual infrequently, either weekly or monthly (see \cref{E-intro:sec:cis}), complicating the data in two ways.
First, infection episodes can go undetected.
Second, the infection episodes' start and end times are known imprecisely.
I explain these issues below and devote the rest of the chapter to the development of a method to tackle  them.

@@

The first complication in the analysis is that infection episodes can be undetected.
Undetected infection episodes are episodes with no associated positive tests.
Therefore, there is no information directly observed about the episode; even the number that occur is latent.
They occur when an episode starts and ends between tests (see \cref{perf-test:fig:truncation}).
Undetected episodes are, on average, shorter than detected episodes; therefore, neglecting them would lead to a biased estimate of the survival distribution.
To see why they are shorter, consider the infection episode in \cref{perf-test:fig:truncation}.
Here two individuals are shown, with identical data collected for each of them (a series of negative
tests).
The first individual was never infected whereas the second individual has an undetected infection. 
The undetected infection episode begins at time 33 and has a 10-day duration.
The individual would have been detected at their next test if their duration had instead been 24 days or more.
The gap between tests in the CIS (up to 28 days, see \cref{E-intro:sec:cis}) is longer than most infections (median 16 days, see \cref{E-ATACCC:sec:results}).
Therefore, it is likely that many infections in the CIS cohort were undetected.
\begin{figure}
  \makebox[\textwidth][c]{\includegraphics[width=\textwidth]{cis-perfect-testing/truncation}}
  \caption[Undetected episodes in CIS data]{%
    Identical data for two individuals, with a series of negative tests.
    Top individual remains uninfected, bottom individual has an undetected infection.
    The shaded area represents when the individual is detectable.
  }
  \label{perf-test:fig:truncation}
\end{figure}

@@

The second complication is that the beginning and end times of detected infection episodes are latent.
Recall, from \cref{E-biology-data:sec:cis-episodes}, that the time episode $j$ starts, $b_j$, is only known to lie in the interval $[l_j^{(b)}, r_j^{(b)}]$ where $l_j^{(b)}$ is the day after the negative test prior to the start of episode $j$ and $r_j^{(b)}$ is the first positive test in episode $j$.
Similarly, for the end of the infection episode, $e_j \in [l_j^{(e)}, r_j^{(e)}]$.
\Cref{perf-test:fig:double-interval-censor} shows this issue graphically.
\begin{figure}
  \makebox[\textwidth][c]{\includegraphics[width=\textwidth]{cis-perfect-testing/double-interval-censor}}
  \caption[Double-interval censoring in CIS data]{%
        The beginning and end of each episode observed in the CIS data is only known up to an interval.
        Demonstrated here by a participant who is recorded negative at time 7 and positive at time 14, bounding the start of the episode; similarly, changing from positive at time 28 to negative at time 56 bounds the end of the episode.
    }
    \label{perf-test:fig:double-interval-censor}
\end{figure}

@@

There are also undetected infections in the CIS (see \cref{perf-test:sec:undetected}).
However, the situation is subtly different from the standard truncation situation encountered in the literature.
In the standard situation, nothing is known about the unrecorded intervals.
But in CIS, individuals in which undetected infections occurred are enrolled in the study.
Therefore, we know the times that they were tested.
This provides information on how long their infections may have been.
I refer to these events as ``undetected'', rather than truncated, to emphasise this difference.

@@

For this conditional likelihood to be valid, they assume this conditioning provides negligible information on the distribution of interest~\citePersonalComms{Nick Jewell}.
This assumption performs well in these settings, when the length of the intervals is fairly constant.
A necessary condition for it to perform well, satisfied when the interval lengths are near-constant, is $p(\text{episode $j$ detected} \mid b_j \in [l_j^{(b)}, r_j^{(b)}]) \approx p(\text{episode $j$ detected})$.
However, the interval lengths change from one to four weeks in the CIS (see \cref{E-intro:sec:cis}).

>>>>

In the CIS, individuals are tested infrequently, either weekly or monthly (see \cref{E-intro:sec:cis}), complicating the data in two ways.
First, infection episodes can go undetected.
Second, the infection episodes' start and end times are known imprecisely.
I explain these issues below and devote the rest of the chapter to the development of a method to tackle  them.

@@

The first complication in the analysis is that infection episodes can be undetected.
Undetected infection episodes are episodes with no associated positive tests.
Therefore, there is no information directly observed about the episode; even the number that occur is latent.
They occur when an episode starts and ends between tests (see \cref{perf-test:fig:truncation}).
Undetected episodes are, on average, shorter than detected episodes; therefore, neglecting them would lead to a biased estimate of the survival distribution.
To see why they are shorter, consider the infection episode in \cref{perf-test:fig:truncation}.
Here two individuals are shown, with identical data collected for each of them (a series of negative
tests).
The first individual was never infected whereas the second individual has an undetected infection. 
The undetected infection episode begins at time 33 and has a 10-day duration.
The individual would have been detected at their next test if their duration had instead been 24 days or more.
The gap between tests in the CIS (up to 28 days, see \cref{E-intro:sec:cis}) is longer than the duration of most infections (median 16 days, see \cref{E-ATACCC:sec:results}).
Therefore, it is likely that many infections in the CIS cohort were undetected.
\begin{figure}
  \makebox[\textwidth][c]{\includegraphics[width=\textwidth]{cis-perfect-testing/truncation}}
  \caption[Undetected episodes in CIS data]{%
    Identical data for two individuals, with a series of negative tests.
    Top individual remains uninfected, bottom individual has an undetected infection.
    The shaded area represents when the individual is detectable.
  }
  \label{perf-test:fig:truncation}
\end{figure}

@@

The second complication is that the beginning and end times of detected infection episodes are latent.
Recall, from \cref{E-biology-data:sec:cis-episodes}, that the time episode $j$ starts, $b_j$, is only known to lie in the interval $[l_j^{(b)}, r_j^{(b)}]$ where $l_j^{(b)}$ is the day after the negative test prior to the start of episode $j$ and $r_j^{(b)}$ is the first positive test in episode $j$.
Similarly, for the end of the infection episode, $e_j \in [l_j^{(e)}, r_j^{(e)}]$.
\Cref{perf-test:fig:double-interval-censor} shows this issue graphically.
\begin{figure}
  \makebox[\textwidth][c]{\includegraphics[width=\textwidth]{cis-perfect-testing/double-interval-censor}}
  \caption[Double-interval censoring in CIS data]{%
        The beginning and end of each episode observed in the CIS data are only known up to an interval.
        Demonstrated here by a participant who is recorded negative at time 7 and positive at time 14, bounding the start of the episode; similarly, changing from positive at time 28 to negative at time 56 bounds the end of the episode.
    }
    \label{perf-test:fig:double-interval-censor}
\end{figure}

@@

There are also undetected infections in the CIS (see \cref{perf-test:sec:undetected}).
However, the situation is subtly different from the standard truncation situation encountered in the literature.
In the standard situation, nothing is known about the unrecorded intervals.
But in CIS, individuals in which undetected infections occurred are enrolled in the study.
Therefore, we know the times that they were tested.
This provides information on how long their infections may have occurred.
I refer to these events as ``undetected'', rather than truncated, to emphasise this difference.

@@

For this conditional likelihood to be valid, they assume this conditioning provides negligible information on the distribution of interest~\citePersonalComms{Nick Jewell}.
This assumption performs well in these settings, when the length of the intervals is fairly constant.
In particular, when the interval lengths are near-constant, $p(\text{episode $j$ detected} \mid b_j \in [l_j^{(b)}, r_j^{(b)}]) \approx p(\text{episode $j$ detected})$.
However, the interval lengths change from one to four weeks in the CIS (see \cref{E-intro:sec:cis}).

### 2024-03-25 introduction.tex

Reproduction numbers are strongly connected to \emph{growth rates}.
Over short periods of time, epidemics tend to grow (or shrink) exponentially, and hence have an exponential growth rate, normally denoted $r$.
For a given disease and population, an analogous growth rate can be defined for each definition of the reproduction number~\autocite{pellisEstimation,paragGrowthRates,wallingaGI}.
The advantage of using the growth rate is that it isn easier to estimate and captures how quickly spread happens in calendar time; however, they do not directly give the strength of intervention required to control the epidemic~\autocite{royalSocietyRnumber}.

@@

The data sources used for outbreak surveillance are often observational in nature.
Therefore, they suffer from biases that can be hard to understand and quantify~\autocite{shadboltChallenges}.
Here I discuss three types of data which are useful for surveillance: incidence-type, prevalence-type, and data from cohort studies.
\emph{Incidence-type} and \emph{prevalence-type} data can directly inform estimation of incidence and transmission; however, this estimation almost always relies on prior knowledge of epidemiological and/or biological quantities.
For example, the time between infection and symptom onset.
These quantities can be estimated from other studies, \eg cohort studies.

@@

Next, are semi-mechanistic models.
These have components of the transmission process in them but do not fully represent the process.
The most common type of semi-mechanistic model is those based on a \emph{renewal equation}~\autocite{bhattSemimechanistic}.
The renewal equation for an infectious disease~\autocite{fraserRenewal}, in the most general form, is:
\begin{align}
    Z(t) &= \int_{0}^{t} \beta(t,\tau) Z(t-\tau) d \tau
    \label{intro:eq:renewal}
\end{align}
where $Z(t)$ is the mean incidence at time $t$ and $\beta(t, \tau)$ is the mean rate an individual infected $\tau$ time units ago infects individuals at calendar time $t$.
Therefore, the renewal equation relates incidence in the past to incidence in the future.
This partially describes the transmission process, hence the name semi-mechanistic.
The effective reproduction number is $\Re = \int_0^\infty \beta(t, \tau) d\tau$.
% A scaled measure of $I(t)$ (\eg hospital admissions or community test results) allows estimating $\beta$.
% The most common implementations are the R packages EpiNow2~\autocite{EpiNow2} and EpiEstim~\autocite{EpiEstim}.
Extensions of \cref{intro:eq:renewal} to track the number of susceptible individuals become full mechanistic models, equivalent to the compartmental models introduced next~\autocite{champredonEquivalence}.
Renewal equations do not feature further in this thesis; for further details, including practical considerations when using them as the basis for inference, see \textcite{nashEstimating,gosticPractical}.

Finally, mechanistic models are the most complex, representing the transmission process in varying levels of detail.
\emph{Compartmental} models, the most common type of mechanistic model, split the population into compartments, each representing a different state of the disease.
The \emph{dynamics} of the epidemic, how it changes over time, are represented by the transitions between compartments (for more details see \cref{E-SEIR:sec:mechanistic-models}).
Simple mechanistic models use a homogeneous population, with individuals differing only due to their disease state, and exponential distributions for the time in each disease state.
More complex models relax these assumptions, in particular to incorporate heterogeneity.
The age structure of the population is important for respiratory diseases because individuals meet individuals of their own age more frequently.
% Mathematically, the dynamics are described by ODEs (ordinary differential equations) or SDEs (stochastic differential equations).
% Unlike renewal equations, there are not widely-used implementations of compartmental models.
% A commonly used frameork is the SEIR (susceptible, exposed, infectious, recovered) model.
% The SEIR model, has four components (susceptible, exposed, infectious, recovered), with the population divided into these components; it is described in detail in \cref{E-SEIR:sec:mechanistic-models}.

@@

Solving \cref{intro:eq:disease-model} for $Z$ is an ill-posed problem.
The forwards error of the solution is large when the ratio of the off-diagonal to diagonal elements is large~\autocite[chapter 8.2]{highamAccuracy}; for this situation, which is when the value of $\max_{i > 0} f(i) / f(0)$ is large.
A large forwards error means that the computed value is far from the true solution~\autocite[6--9]{highamAccuracy}; this is avoided by either using an approach which gives a large $f(0)$ or smoothing $Z_t$.

@@

For large parts of the pandemic, community testing was widely available in the UK.
However, it did not test a random or even representative sample of the population.
The number of tests conducted varied over time and space, for a variety of known and unknown reasons.
These include: the policy on who is eligible for testing, the availability of tests, and the willingness of individuals to be tested.
Therefore, fluctuations in the number of positive tests may not reflect changes in the true incidence of the disease.
The large number of tests conducted allowed detailed analyses to be performed, including stratification by small geographic areas or age groups, although this does not solve the issue of the biases present.

Severe events counts are generally a reliable measure of the burden of the disease.
The criteria for being a severe event are normally well-defined and generally well-recorded in the UK context.
Therefore, these data are subject to fewer biases than community testing.
However, SARS-CoV-2 rarely leads to severe events in younger individuals, meaning severe event counts are insensitive to changes in transmission only occurring in younger individuals.

@@

The extraordinary impact of SARS-CoV-2 led the UK Government to fund two large prevalence surveys in representative samples of the general population.
These were not subject to the biases of community testing and observed the whole population, unlike severe events, although suffered from non-response bias.
Therefore, some modelling is required to borrow information across strata or time to provide stable estimates, as well as to correct for non-response bias (see \cref{E-intro:sec:cis}).
Despite this, they are considered the gold-standard metric.

>>>>

Reproduction numbers are strongly connected to \emph{growth rates}.
Over short periods of time, epidemics tend to grow (or shrink) exponentially, and hence have an exponential growth rate, normally denoted $r$.
For a given disease and population, an analogous growth rate can be defined for each definition of the reproduction number~\autocite{pellisEstimation,paragGrowthRates,wallingaGI}.
The advantages of using the growth rates are that they are easier to estimate and capture how quickly spread happens in calendar time; however, they do not directly give the strength of intervention required to control the epidemic~\autocite{royalSocietyRnumber}.

@@

Data sources used for outbreak surveillance are often observational in nature.
Therefore, they suffer from biases that can be hard to understand and quantify~\autocite{shadboltChallenges}.
Here I discuss three types of data which are useful for surveillance: incidence-type, prevalence-type, and data from cohort studies.
\emph{Incidence-type} and \emph{prevalence-type} data can inform estimation of incidence and transmission; however, this estimation almost always relies on prior knowledge of epidemiological and/or biological quantities.
For example, the time between infection and symptom onset.
These quantities can be estimated from other studies, \eg cohort studies.

@@

Next, are semi-mechanistic models.
These have components of the transmission process in them but do not fully represent the process.
A common type of semi-mechanistic model is those based on a \emph{renewal equation}~\autocite{bhattSemimechanistic}.
The renewal equation for an infectious disease~\autocite{fraserRenewal}, in the most general form, is:
\begin{align}
    Z(t) &= \int_{0}^{t} \beta(t,\tau) Z(t-\tau) d \tau
    \label{intro:eq:renewal}
\end{align}
where $Z(t)$ is the mean incidence at time $t$ and $\beta(t, \tau)$ is the mean rate at which an individual infected $\tau$ time units ago infects individuals at calendar time $t$.
Therefore, the renewal equation relates incidence in the past to incidence in the future.
This partially describes the transmission process, hence the name semi-mechanistic.
The effective reproduction number is $\Re = \int_0^\infty \beta(t, \tau) d\tau$.
% A scaled measure of $I(t)$ (\eg hospital admissions or community test results) allows estimating $\beta$.
% The most common implementations are the R packages EpiNow2~\autocite{EpiNow2} and EpiEstim~\autocite{EpiEstim}.
Extensions of \cref{intro:eq:renewal} that track the number of susceptible individuals become full mechanistic models, equivalent to the compartmental models introduced next~\autocite{champredonEquivalence}.
Renewal equations do not feature further in this thesis; for further details, including practical considerations when using them as the basis for inference, see \textcite{nashEstimating,gosticPractical}.

Finally, mechanistic models are the most complex models, representing the transmission process in varying levels of detail.
\emph{Compartmental} models, the most common type of mechanistic model, split the population into compartments, each representing a different state of the disease.
The \emph{dynamics} of the epidemic, how it changes over time, are represented by the transitions between compartments (for more details see \cref{E-SEIR:sec:mechanistic-models}).
Simple mechanistic models use a homogeneous population, with individuals differing only due to their disease state, and exponential distributions for the time in each disease state.
More complex models relax these assumptions, in particular to incorporate heterogeneity.
The age structure of the population is important for respiratory diseases because individuals meet individuals of their own age more frequently.
% Mathematically, the dynamics are described by ODEs (ordinary differential equations) or SDEs (stochastic differential equations).
% Unlike renewal equations, there are not widely-used implementations of compartmental models.
% A commonly used frameork is the SEIR (susceptible, exposed, infectious, recovered) model.
% The SEIR model, has four components (susceptible, exposed, infectious, recovered), with the population divided into these components; it is described in detail in \cref{E-SEIR:sec:mechanistic-models}.

@@

Solving \cref{intro:eq:disease-model} for $Z$ is an ill-posed problem.
The forwards error of the solution is large when $\max_{i > 0} f(i) / f(0)$~\autocite[chapter 8.2]{highamAccuracy}.
A large forwards error means that the computed value is far from the true solution~\autocite[6--9]{highamAccuracy}; this is avoided by either using an approach which gives a large $f(0)$ or smoothing $Z_t$.

@@

For large parts of the pandemic, community testing was widely available in the UK.
However, it did not test a representative sample of the population.
The number of tests conducted varied over time and space, for a variety of known and unknown reasons.
These include: the policy on who is eligible for testing, the availability of tests, and the willingness of individuals to be tested.
Therefore, fluctuations in the number of positive tests may not reflect changes in the true incidence of the disease.
The large number of tests conducted allowed detailed analyses to be performed, including stratification by small geographic areas or age groups, although this does not solve the issue of the biases present.

Severe events counts are generally a more reliable measure of the burden of the disease.
The criteria for being a severe event are normally well-defined and generally well-recorded in the UK context.
Therefore, these data are subject to fewer biases than community testing.
However, SARS-CoV-2 rarely leads to severe events in younger individuals, meaning severe event counts are insensitive to changes in transmission only occurring in younger individuals.

@@

The extraordinary impact of SARS-CoV-2 led the UK Government to fund two large prevalence surveys in representative samples of the general population.
These were designed to avoid the biases of community testing and observe the whole population.
Therefore, some modelling is required to borrow information across strata or time to provide stable estimates, as well as to correct for non-response bias (see \cref{E-intro:sec:cis}).
Despite this, they are considered the gold-standard metric.

### 2024-03-25 thesis.tex

\abstract{%
    Surveillance is a crucial component to a pandemic response, informing public health and wider policy.
    Unfortunately, it is often subject to reporting and behavioural biases, reducing its robustness.
    % During the SARS-CoV-2 epidemic, the UK government funded two large-scale prevalence surveys with samples recruited randomly from the general population.
    % One of these, the CIS (Coronavirus Infection Survey), was a longitudinal survey run from April 2020 to March 2023.
    The CIS (Coronavirus Infection Survey) was a large-scale prevalence survey, with participants recruited randomly from the general population, therefore avoiding these issues.

@@

    This relies on knowledge of the duration of positivity (how long individuals test positive for).
    Therefore, I first consider how to estimate the distribution of this duration in the general population.
    I extend a pre-existing model of how an individual's viral load changes over time to estimate the duration of positivity.
    % I take two complementary approaches using different datasets.
    % This model is fitted to data from a longitudinal study of contacts of individuals with confirmed SARS-CoV-2 infection.
    % In \cref{ATACCC}, I extend a pre-existing model of how an individual's viral load changes over time to estimate the duration of positivity.
    % This model is fitted to data from the ATACCC (Assessment of Transmission and Contagiousness of COVID-19 in Contacts) study, a longitudinal study of viral load in contacts of individuals with confirmed SARS-CoV-2 infection.
    I use this detailed model to exploit frequent (daily) sampling of individuals in a cohort of contact traced individuals.
    However, the cohort was small and follow-up short meaning that the tail of the duration distribution is poorly estimated.

    % Therefore, in \cref{perf-test,imperf-test}, I combine the information from ATACCC with the larger CIS dataset.
    % I analyse the CIS in a survival analysis framework.
    To overcome this issue, I combine the information with the larger and longer CIS dataset.
    The CIS data poses several challenges because it is double interval censored and some individuals in the cohort have infections that are never detected.
    I apply a survival analysis framework to analyse this data, which had previously only been used in simpler settings.
    In addition, I develop novel methodology to incorporate false negatives into the framework.
    % The results showed that false negatives in the CIS data needed to be included in the model.
    % In \cref{imperf-test} I develop a novel methodology to do so.
    This results in an estimate of the full duration distribution in the general population, not previously available.

    Finally, I use this estimate to estimate the incidence and transmission of SARS-CoV-2 in England.
    I fully propagate the sampling error, accurately include the shape of the duration distribution, and estimate the quantities by age; previous work lacked at least one of these.
    Furthermore, I fit to only the CIS data, showing that such a survey is adequate to estimate these quantities.
    % without requiring an evidence synthesis approach previously used.
}

@@

    In addition to the supervision and advice mentioned, several people contributed code or results to this work.
    Professor A.\ Sarah Walker created the episodes dataset in \cref{biology-data:sec:cis-episodes} and performed the initial survival analysis I use for a simulation study \cref{perf-test:sec:simulation-study}.
    Dr Koen Pouwels provided code used as the basis of the positivity model in \cref{biology-data:sec:MRP}.
    Dr Edwin van Leeuwen created the contact matrices and Dr Sanmitra Ghosh provided code for the MCMC sampler used in \cref{E-SEIR}.
    Thank you all for your contributions.

@@

    Several aspects of my work were done to support the ongoing response.
    In particular, an earlier version of the backcalculation work was adopted by the Office for National Statistics to estimate incidence, published in their weekly online reports~\autocite{cisMethodsONS}.
    This method was also used for the analysis of several respiratory diseases over winter 2022/3~\autocite{dietzSARSCoV2final}.

>>>>

\abstract{%
    Surveillance is a crucial component of a pandemic response, informing public health and wider policy.
    Unfortunately, it is often subject to reporting and behavioural biases, reducing its robustness.
    % During the SARS-CoV-2 epidemic, the UK government funded two large-scale prevalence surveys with samples recruited randomly from the general population.
    % One of these, the CIS (Coronavirus Infection Survey), was a longitudinal survey run from April 2020 to March 2023.
    The CIS (Coronavirus Infection Survey) was a large-scale prevalence survey, with participants recruited randomly from the general population, thereby avoiding these issues.

@@

    This relies on knowledge of the duration of positivity (how long individuals test positive for).
    Therefore, I first consider how to estimate the distribution of this duration in the general population.
    I extend a pre-existing model of how an individual's viral load changes over time to estimate the duration of positivity.
    % I take two complementary approaches using different datasets.
    % This model is fitted to data from a longitudinal study of contacts of individuals with confirmed SARS-CoV-2 infection.
    % In \cref{ATACCC}, I extend a pre-existing model of how an individual's viral load changes over time to estimate the duration of positivity.
    % This model is fitted to data from the ATACCC (Assessment of Transmission and Contagiousness of COVID-19 in Contacts) study, a longitudinal study of viral load in contacts of individuals with confirmed SARS-CoV-2 infection.
    I use this detailed model to exploit frequent (daily) sampling of individuals in a cohort of contact traced individuals.
    However, the cohort was small and the follow-up time short meaning that the tail of the duration distribution is poorly estimated.

    % Therefore, in \cref{perf-test,imperf-test}, I combine the information from ATACCC with the larger CIS dataset.
    % I analyse the CIS in a survival analysis framework.
    To overcome this issue, I combine the information with the larger and longer CIS dataset.
    The CIS data poses several challenges because it is doubly interval censored and some individuals in the cohort have infections that are never detected.
    I apply a survival analysis framework to analyse this data, which had previously only been used in simpler settings.
    In addition, I develop novel methodology to incorporate false negatives into the framework.
    % The results showed that false negatives in the CIS data needed to be included in the model.
    % In \cref{imperf-test} I develop a novel methodology to do so.
    This results in an estimate of the full duration distribution in the general population, not previously available.

    Finally, I use this estimate to infer the incidence and transmission of SARS-CoV-2 in England from the CIS.
    I fully propagate the sampling error, accurately include the shape of the duration distribution, and estimate the quantities by age; previous work lacked at least one of these.
    Furthermore, I fit to only the CIS data, showing that such a survey is adequate to estimate these quantities.
    % without requiring an evidence synthesis approach previously used.
}

@@

    In addition to the supervision and advice mentioned, several people contributed code or results to this work.
    Professor A.\ Sarah Walker created the episodes dataset in \cref{biology-data:sec:cis-episodes} and performed the initial survival analysis I use for a simulation study \cref{perf-test:sec:simulation-study}.
    Dr Koen Pouwels provided code I use as the basis of the positivity model in \cref{biology-data:sec:MRP}.
    Dr Edwin van Leeuwen created the contact matrices and Dr Sanmitra Ghosh provided code for the MCMC sampler I use in \cref{E-SEIR}.
    Thank you all for your contributions.

@@

    Several aspects of the work in this thesis were completed in support of the ongoing response.
    In particular, an earlier version of the backcalculation work was adopted by the Office for National Statistics to estimate incidence, published in their weekly online reports~\autocite{cisMethodsONS}.
    This method was also used for the analysis of several respiratory diseases over winter 2022/3~\autocite{dietzSARSCoV2final}.
